---
title: "w2 exercises"
author: "Balazs Szabo"
date: "2/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE, root.dir = '.')
```
# Libraries
```{r}
library(tidyverse)

```
#3.1
higher taxes are supported by 50% of respondents
aged 18–29, 60% of respondents aged 30–44, 40% of respondents aged 45–64, and 30% of
respondents aged 65 and up.

200 respondents aged 18–29, 250 aged 30–44, 300 aged 45–64,
and 250 aged 65+.
```{r}
age1 <- 200
age2 <- 250
age3 <- 300
age4 <- 250

(0.5*age1+0.6*age2+0.4*age3+0.3*age4)/(age1+age2+age3+age4)

```
44.5% supports higher taxes

#3.3 
```{r}
set.seed(1666)

norm1 <- rnorm(100,5,2)
norm2 <- rnorm(100,10,10)
norm3 <- rnorm(100,5,10)

Df_1 <- data.frame(norm1) 
Df_1 <- Df_1 %>% 
  mutate(name = "norm1") %>% 
  mutate(value = norm1) %>% 
  dplyr::select(!norm1)

Df_2 <- data.frame(norm2)
Df_2 <- Df_2 %>% 
  mutate(name = "norm2") %>% 
  mutate(value = norm2) %>% 
  dplyr::select(!norm2)

Df_3 <- data.frame(norm3)
Df_3 <- Df_3 %>% 
  mutate(name = "norm3") %>% 
  mutate(value = norm3) %>% 
  dplyr::select(!norm3)

Df_4 <- rbind(Df_1,Df_2,Df_3)

Df_4 %>% 
  ggplot(aes(x=value, fill=name)) +
  geom_density(alpha=0.5)
```
#3.4
```{r}
set.seed(1666)

df_poison <- data.frame(rpois(20, 3.5))

df_poison <- df_poison %>% 
  mutate(asd=rpois.20..3.5.)

df_poison %>% 
  ggplot(aes(x=asd)) +
  geom_bar()
```
#3.5
```{r}
df_bin <- data.frame(rbinom(20,20,0.3))
df_bin %>% 
  ggplot(aes(x=rbinom.20..20..0.3.)) +
  geom_bar()

```
#3.6
##A
```{r}
#figuring out how to solve this one case
set.seed(1666)
library(truncnorm)
test<- rtruncnorm(n = 100,a = 0,b = 50,mean = 35,sd = 10)

testmean <- vector(mode="numeric", length=length(test))
testmean[1:length(testmean)] <- mean(test)

mean(36.56+1.793*test)
sd(36.56+1.789*test)

mean(test)
sd(test)

newssd <- 15/(sqrt((sum((test-testmean)^2)/(length(test)))))

100-mean(newssd*test)

```
```{r}
#writing a function to solve most? cases
linear_transform_it <- function(dist, newsd, newmean){
  #makes a vector with the same length of the original one but empty
  distmean <- vector(mode="numeric", length=length(dist))
  #fills that empty vector with the mean value
  distmean[1:length(distmean)] <- mean(dist)
  # this part uses the standard deviation * X = new standard deviation
  # formula to solve for X
  multby_sd <- newsd/(sqrt((sum((dist-distmean)^2)/(length(dist)))))
  #this part multiplies the original vector by X,and subtracts it from the new_mean to
  # figure out which number should be added to the linear transformation
  add_to_mean <- newmean-mean(multby_sd*test)
  
  
  
  sprintf("%s+%s*original vector",
          round(add_to_mean,digits = 4),round(multby_sd, digits=4))
  
  
}
```

```{r}
linear_transform_it(test,15,100)
```
```{r}
mean(36.3742+1.793*test)
sd(36.3742+1.793*test)

test2 <- 36.3742+1.793*test

```
##B
```{r}
test2 <- 36.3742+1.793*test
range(test)
range(test2)

36.3742+1.793*12.69608
36.3742+1.793*49.92444
```
The new values range from a+bx1, where x is the smallest original value to a+bxn, where xn is the largest original value

##C

```{r}
plot(x=test,y=test2, type='l',col='pink',xlab='x',ylab='y')
text(x=test[20],y=test2[50],labels='a+bx',col='black')
```


#4.1  
```{r}
se_t <- sqrt(0.5*0.5/500)

se_c <- sqrt(0.4*0.6/500)

estimate <- 0.5-0.4

se_estimate <- sqrt(se_t^2+se_c^2)

estimate
se_estimate
```
The estimated difference between the treatment and control group is 0.1, with a standard error of 0.03130

#4.2

If p is near 0.5 we can approximate standard error as 0.5/sqrt(n).

sqrt(0.5/sqrt(n)^2 + 0.5/sqrt(n)^2) < 0.05

this can be rearranged and solve for n


```{r}

n <- (2*0.5^2)/0.05^2

n*2

```
at least 400 samples are needed

#4.3

```{r}

se_w <- sqrt(0.3*0.7/20)

se_b <- sqrt(0.4*0.6/20)

estimate4_3 <- 0.4-0.3

se_estimate4_3 <- sqrt(se_w^2+se_b^2)

estimate4_3
se_estimate4_3
# calculate t score
t_stats4_3 <- estimate4_3/se_estimate4_3
t_stats4_3

pt(t_stats4_3,19)

```
The probability is around 74%

#4.4


```{r}

n2 <- 20

while( 0.95 > pt((0.4-0.3)/sqrt((sqrt(0.3*0.7/n2))^2+(sqrt(0.4*0.6/n2))^2),n2-1))
{


n2 <- n2+1



}

print(n2)

```
with an N of 124 there is a 95% probability of distinguishing the shooters



