---
title: "Custom_Functions"
author: "Balazs"
date: "10/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE, root.dir = '.')
library(tidyverse)
library(car)
```
  
# Intro  
Hi guys, I thought these two functions might be useful to you too  
The "is_this_normal" will take a dataset and tell you if it is distributed normally.  
The "quick_hist" will take a dataset and make a histogram about it. This only has Tidyverse elements, so it works with %>% 


```{r}
is_this_normal<- function(dataframe, var1, binwidth = 30){
  ## a function, which combine "is_this_normal" and "quick_hist"
 
   ## a function that helps decide if a data set is normally distributed  
  library(pastecs)
  library(tidyverse)
  
## very important step!!!!! -- > this makes it work with pipes
 DF1 <- dataframe %>% 
   mutate(var1_1 = {{var1}} ) %>% 
   select(var1_1)
 
##  A vector with the important numeric data
Table1 <- stat.desc(DF1$var1_1, basic = FALSE, norm = TRUE) 

## Printing out results
  
  if ((Table1[c("skew.2SE")]<1) & (Table1[c("kurt.2SE")]<1) & (Table1[c("normtest.p")]>0.05)) 
    {
    print("Null hypothesis wasn't disproved, kurtosis ans skewness are acceptable")
  } else 
    {
    print("Not normally distributed")
    }
  print(Table1[c("skewness","kurtosis","skew.2SE","kurt.2SE","normtest.p")])
 
 
## a function to get a quick histogram of a data set, along with the mean, SD and number of samples (Also works with %>% )

  
## very important step!!!!! -- > this makes it work with pipes
 DF2 <- dataframe %>% 
   mutate(var1_2 = {{var1}} ) %>% 
   select(var1_2)
 
 #name of the plot
 STRNamevar <- deparse(substitute(var1))

## a Quick summary of the mean, Standard deviation and sample size, not that formative, but might be enough in some cases
DF3 <- DF2 %>% 
  select(var1_2) %>% 
  summarize( Mean=mean(var1_2), StandardDeviation=sd(var1_2), NumberOfSamples=n() )


## A histogram showing the data with the mean as a dotted line
Plot1 <- DF2 %>% 
  ggplot(aes(x = var1_2))+
  geom_histogram(aes(y=..density..),color="Black", fill= "Red", alpha=0.5, binwidth = binwidth) +
  geom_vline(data=DF2, aes(xintercept=mean(var1_2)),
             linetype="dashed") +
  stat_function(fun = dnorm,
                args = list(mean = mean(DF2$var1_2, na.rm = TRUE),
                sd = sd(DF2$var1_2, na.rm = TRUE)),
                colour= "black", size = 1) +
ggtitle(STRNamevar)

## QQ plot 
Plot2 <- DF2 %>% 
  ggplot(aes(sample = var1_2))+
stat_qq() +
  stat_qq_line() +
  ggtitle(STRNamevar)

  
## Printing out results
  print(DF3)
  print(Plot1)
  print(Plot2)
 
}

```

```{r}

no_outliers <- function(dataframe,var1){
  
## A function for removing outliers from the data
  library(tidyverse)
  
  ## very important step!!!!! -- > this makes it work with pipes
 DF1 <- dataframe %>% 
   mutate(var1_1 = {{var1}} ) %>% 
   select(var1_1)
 
  ## defining quantiles
  Q <- quantile(DF1$var1_1, probs=c(.25, .75), na.rm = FALSE)
  iqr <- iqr <- IQR(DF1$var1_1)
  ## explanation to what an upper and lower range is
  up <-  Q[2]+1.5*iqr # Upper Range  
  low <- Q[1]-1.5*iqr # Lower Range
  
  ##new dataframe without ouliers
  without_outliers<- subset(DF1, var1_1 > (Q[1] - 1.5*iqr) & var1_1 < (Q[2]+1.5*iqr))
  
 ##making it pipe  output compatible
   STRNamevar <- c(deparse(substitute( var1 )))
   DFOut <- without_outliers %>% 
   mutate( !!STRNamevar[1] := var1_1 ) %>% 
     select(!var1_1)
  
  
   
  # rename(without_outliers, STRNamevar = var1_1)
   return(DFOut)
   
}

```


# In Action:  
I got a dataset with all the Pokemon data and tried to find something with a normal distribution
```{r}
DF1_all_data <- read_csv("pokemonsdataset.csv")
```


```{r}
trans_to_norm_log<- function(dataframe, var1){
  #positive skew logarithmic transformations function for non normally distributed data
  
  ## very important step!!!!! -- > this makes it work with pipes
 DF1 <- dataframe %>% 
   mutate(var1_1 = {{var1}} ) %>% 
   select(var1_1)

  #transforming
DFTrans <- DF1 %>% 
  mutate(logvar1 = log( var1_1 ))  %>% 
  select(logvar1)
 

##making it pipe  output compatible
   STRNamevar <- c(deparse(substitute( var1 )))
   DFOut <- DFTrans %>% 
   mutate( !!STRNamevar[1] := logvar1 ) %>% 
     select(!logvar1)

return(DFOut)
   
}
```

```{r}
trans_to_norm_sqrt <- function(dataframe, var1){
  
  #positive skew squared transformations function for non normally distributed data
  
  ## very important step!!!!! -- > this makes it work with pipes
 DF1 <- dataframe %>% 
   mutate(var1_1 = {{var1}} ) %>% 
   select(var1_1)

  #transforming
DFTrans <- DF1 %>% 
  mutate(sqrtvar1 = sqrt( var1_1 ))  %>% 
  select(sqrtvar1)
 

##making it pipe  output compatible
   STRNamevar <- c(deparse(substitute( var1 )))
   DFOut <- DFTrans %>% 
   mutate( !!STRNamevar[1] := sqrtvar1 ) %>% 
     select(!sqrtvar1)

return(DFOut)
  
}

```

```{r}
trans_to_norm_recip <- function(dataframe, var1){
  
  #positive skew squared transformations function for non normally distributed data
  
  ## very important step!!!!! -- > this makes it work with pipes
 DF1 <- dataframe %>% 
   mutate(var1_1 = {{var1}} ) %>% 
   select(var1_1)

  #transforming
DFTrans <- DF1 %>% 
  mutate(recipvar1 = 1/ var1_1 )  %>% 
  select(recipvar1)
 

##making it pipe  output compatible
   STRNamevar <- c(deparse(substitute( var1 )))
   DFOut <- DFTrans %>% 
   mutate( !!STRNamevar[1] := recipvar1 ) %>% 
     select(!recipvar1)

return(DFOut)
  
}
```


```{r}

leveneTest(DF1_all_data$Attack,DF1_all_data$`Primary Type`)



```

